---
title: 'Homework Assignment #2'
author: "Math 437 - Modern Data Analysis"
date: "Due February 24, 2023"
output: pdf_document
---

# Instructions

You should submit either two or three files:

1. You should write your solutions to the Simulation and Applied Problems in this R Markdown file and submit the (.Rmd) file.
2. You should knit the final solution file to pdf and submit the pdf. If you are having trouble getting code chunks to run, add `eval = FALSE` to the chunks that do not run. If you are having trouble getting R Studio to play nice with your LaTeX distribution, I will begrudgingly accept an HTML file instead. 
3. Solutions to the Key Terms and Conceptual Problems can be submitted in a separate Word or pdf file or included in the same files as your solutions to the Simulation and Applied Problems.

This homework assignment is worth a total of **40 points**.

# Key Terms (5 pts)

Read Chapter 13 of Introduction to Statistical Learning, Second Edition. Based on your reading, answer the following questions.

1. What is a *p-value*? What is the difference between a one-sided and a two-sided p-value?
A $p-value$ is defined as the probability of observing a test statistic equal to or more extreme than the observed statistic, under the assumption that H_0 is true.

2. In traditional NHST-style significance testing, what are the two possible decisions? When do we make each decision?
The two possible decisions are to reject the null hypothesis or fail to reject it. We make each decision based on what our p-value is. A smaller p-value indicates that there is evidence against H_0, thus provides evidence against it. 

3. What is the difference between a *Type I Error* and a *Type II Error*?
A $Type I Error$ is when we reject H_0 when it is actually true. A $Type II Error$ is when we do not reject H_0 even though it is false. 

4. Briefly explain why it is necessary to adjust the significance level (or equivalently, the p-values) when testing a large number of null hypotheses.
It is necessary to adjust the significance level when testing a large number of null hypotheses because there is a possibility of rejecting a lot of true null hypotheses, leading to a great number of Type I Errors.

5. Compare and contrast the *Family-Wise Error Rate* (FWER) and the *False Discovery Rate* (FDR).
`FWER` is the probability of making at least one Type I Error while `FDR` is a ratio of false positives to total positives to ensure that we are rejecting as many null hypotheses as possible while guaranteeing that the a certain percentage of those rejected null hypotheses are false positives. It is impossible to control the `FDR` but we can control the `FWER`. #################

6. Compare and contrast the *Bonferroni Method* and *Holm's Step-Down Method* for controlling the FWER.
The `Bonferroni method` refers to adjusting the alpha in FWER to control the probability of performing a Type I Error. `Holm's method` also controls the FWER, but it is adjusted to result in fewer Type II Errors. Holm's method is less conservative to Bonferroni's method, and uniformly more powerful. In Holm's method, the threshold depends on the values of all m p-values, while Bonferri's method depends on the p-value lower than the ratio of alpha over m p-values.

7. Why do we prefer to use *Tukey's Method* or *Scheffe's Method* to control the FWER? In what conditions is it appropriate to use those methods instead of the Bonferroni or Holm methods?
There are certain very specific settings that we prefer to use $Tukey's Method$ or $Scheffe's Method$.
$Tukey's Method$ allows us to control the FWER at some level alpha while rejecting all null hypotheses where the p-value falls below the alpha. ################

8. Briefly describe the *Benjamini-Hochberg* procedure for controlling the FDR.
The `Benjamini-Hochberg` procedures ensures that, on average, no more than a fraction q of the rejected null hypotheses are false positives. First, we determine what level we want to control the FDR, known otherwise as q. Then we compute the p-values #####

9. What is/are the major assumption(s) of a *permutation test*? What is the general procedure for obtaining the null distribution of a test statistic using a permutation test?
########

10. When is it useful/recommended to use a permutation testing approach as opposed to a traditional theory-based approach?
#########

# Conceptual Problems

## Conceptual Problem 1 (5 pts) 

Textbook Exercise 13.7.6 #######
(a) How many false positives, false negatives, true positives, true
negatives, Type I errors, and Type II errors result from applying
the Bonferroni procedure to control the FWER at level $\alpha = 0.05$?
**refer to Page 559 in ISLR2 textbook**

(b) How many false positives, false negatives, true positives, true
negatives, Type I errors, and Type II errors result from applying
the Holm procedure to control the FWER at level $\alpha = 0.05$?


(c) What is the false discovery rate associated with using the Bonferroni procedure to control the FWER at level $\alpha = 0.05$?


(d) What is the false discovery rate associated with using the Holm
procedure to control the FWER at level $\alpha = 0.05$?


(e) How would the answers to (a) and (c) change if we instead used
the Bonferroni procedure to control the FWER at level $\alpha = 0.001$?


## Conceptual Problem 2 (2 pts)

Suppose that we test $m = 1000$ independent null hypotheses, of which 10% are true, at significance level $\alpha = 0.05$ and achieve a false discovery rate of $q = 0.20$. Construct a table following Table 13.2 in the textbook, identifying the appropriate values of $V$, $S$, $U$, $W$, and $R$ in this situation.
#####


## Conceptual Problem 3 (3 pts)

Suppose that we test $m = 1000$ independent null hypotheses, of which an unknown number $m_0$ are true, at significance level $\alpha = 0.05$. Suppose that each test also has a power of 0.80. Find and plot the false discovery rate as a function of $m_0$.
#####


## Conceptual Problem 4 (2.5 pts)

Textbook Exercise 5.4.2 parts (a), (b), and (c).

We will now derive the probability that a given observation is part
of a bootstrap sample. Suppose that we obtain a bootstrap sample
from a set of n observations.

(a) What is the probability that the first bootstrap observation is
not the jth observation from the original sample? Justify your
answer.
The probability that the first boothstrap observation is NOT the jth observation is 1 - 1/n because 1/n is if the jth observation is selected as the first bootstrap observation.

(b) What is the probability that the second bootstrap observation
is not the jth observation from the original sample?
Since bootstrapping is done with replacement, the probability that the second boostrap observation is not the jth observation is also 1 - 1/n.

(c) Argue that the probability that the jth observation is not in the
bootstrap sample is (1 − 1/n)^n.
The probability that the jth observation not in the boostrap sample is not (1-1/n)^n because we are boostrapping without replacement which means the probability would alwayas be 1-1/n.

# Simulation Problems

## Simulation Problem 1 (Code: 1 pt; Explanation: 0.5 pts)

Textbook Exercise 5.4.2 parts (e), (g), and (h). For part (g), you should create a line plot (using either `plot` with argument `type = "l"` or `geom_line`). Then, to make clearer what you should be commenting on, find the limit as $n \rightarrow \infty$ of the probability that the $j^{th}$ observation is in your bootstrap sample and add a horizontal red line (using `abline` or `geom_hline`) at that value. (Hint: the limit as $n \rightarrow \infty$ of the expression in part (c) is well-known and easily found on the Internet.)

(e) When n = 100, what is the probability that the jth observation is in the bootstrap sample?
The probability when n = 100 is (1 - (1/100)^100 = 0.634

(g) Create a plot that displays, for each integer value of n from 1 to 100, 000, the probability that 
the jth observation is in the bootstrap sample. Comment on what you observe.
```{r}

```
##########

(h) We will now investigate numerically the probability that a bootstrap sample of size n = 100 
contains the jth observation. Here j = 4. We repeatedly create bootstrap samples, and each time
we record whether or not the fourth observation is contained in the bootstrap sample.
###########

## Simulation Problem 2 (Code: 1.5 pts; Explanation: 3.5 pts)

Copy the *functions* you created in the Bootstrap Confidence Intervals class activity as well as Simulation Parts 3, 4, and 5.

Write a brief summary of what you learned from the activity. Make sure to address the following questions:

- Are theory-based methods *guaranteed* to achieve the appropriate coverage? What about bootstrap-based methods?
- Which of the four methods appear to be range-preserving even in a "worst-case scenario"? 
- When and why would a bootstrap method be useful to obtain a confidence interval even if it doesn't achieve the appropriate coverage?



# Applied Problems

## Applied Problem 1 (Code: 4 pts; Explanation: 2 pts)

Using the `dplyr` package, subset the `mpg` dataset from the `ggplot2` package to include only the cars from 2008 that are minivans, pickups, or SUVs (`%in%` is a useful replacement for `==` when trying to match to more than one possibility). Using this new dataset, determine which of the following statements is/are true, using an $\alpha = 0.10$ significance level/family-wise error rate or a $q = 0.10$ false discovery rate:

1. There is a significant difference in highway gas mileage between minivans and SUVs.
2. There is a significant difference in highway gas mileage between pickups and SUVs.
3. There is a significant difference in highway gas mileage between minivans and pickups.

Use the following methods.

(a) Three two-sample t-tests with no adjustments for multiple testing. Store all three p-values in a single vector so that you can use the `p.adjust` function in later parts.
(b) Three two-sample t-tests followed by Bonferroni's method. 
(c) Three two-sample t-tests followed by Holm's step-down method.
(d) A one-way ANOVA followed by Tukey's method.
(e) Three two-sample t-tests followed by the Benjamini-Hochberg (BH) method.

Compare and contrast your results.

## Applied Problem 2 (Code: 1 pt; Explanation: 1 pt)

Use a one-way ANOVA followed by Scheffe's method (`ScheffeTest` in the DescTools package) to determine whether the following statement is true at the $\alpha = 0.10$ significance level:

There is a significant difference in highway gas mileage between pickups and non-pickups (SUVs and minivans).

## Applied Problem 3 (Code: 5 pts; Explanation: 3 pts)

Textbook Exercise 5.4.9.

We will now consider the Boston housing data set, from the ISLR2 library.
(a) Based on this data set, provide an estimate for the population mean of medv. Call this estimate ˆµ.
```{r}
attach(Boston)
mu = mean(Boston$medv)
print(mu)
```

(b) Provide an estimate of the standard error of ˆµ. Interpret this result.
Hint: We can compute the standard error of the sample mean by dividing the sample standard deviation by the square root of the number of observations.
```{r}
sd(Boston$medv)/sqrt(nrow(Boston))
```

(c) Now estimate the standard error of ˆµ using the bootstrap. How does this compare to your answer from (b)?  ######################
```{r}
boot.fn = function(data, index) {
  return(mean(data[index]))
}

BS = boot(Boston$medv, boot.fn, 1000)
print(BS)
```
The answer in part (b) is 0.40886 which is slightly less than the standard error from part (c) of 0.41318.

(d) Based on your bootstrap estimate from (c), provide a 95 % confidence interval for the mean of medv. Compare it to the results obtained using t.test(Boston$medv).
Hint: You can approximate a 95 % confidence interval using the formula [ˆµ − 2SE(ˆµ), µˆ + 2SE(ˆµ)].
```{r}
conf.int = c(BS$t0 - (2 * 0.3994119), BS$t0 + (2  * 0.3994119))

t.test(Boston$medv)
```


(e) Based on this data set, provide an estimate, ˆµmed, for the median
value of medv in the population.
```{r}
mu_med = median(Boston$medv)
print(mu_med)
```


(f) We now would like to estimate the standard error of ˆµmed. Unfortunately, there is no simple formula for computing the standard error of the median. Instead, estimate the standard error of the median using the bootstrap. Comment on your findings.
```{r}
boot.median.fn = function(data, index) {
  return(median(data[index]))
}

BS.median = boot(Boston$medv, boot.median.fn, 10000)
print(BS.median)
```

(g) Based on this data set, provide an estimate for the tenth percentile of medv in Boston census tracts. Call this quantity ˆµ0.1. (You can use the quantile() function.)
```{r}
mu0.1 = quantile(Boston$medv, probs=c(.1))
print(mu0.1)
```


(h) Use the bootstrap to estimate the standard error of ˆµ0.1. Comment on your findings.
```{r}
boot.quant.ten.fn = function(data, index) {
  return(quantile(data[index], probs=c(.1)))
}

boot.quant.ten = boot(Boston$medv, boot.quant.ten.fn, 10000)
print(boot.quant.ten)
```

